---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# aMSE

<!-- badges: start -->
<!-- badges: end -->

This packages up a new Abalone Management Strategy Evaluation framework. It has a novel structure that is based around the spatial ideas of populations within spatial management units (SMUs), within a region:

* region - highest geogrpahical level. Is simply the totality of the spatial management units.

* SMU - spatial management units. In Tasmania these would currently be the classical statistical blocks.

* population - literally a population. These are the active components within the simulation framework. The dynamics of the simulation are based around the populations, although, with positive larval dispersal (the default) there is some dependency of neighboiuring populations on each other.


This conceptual structure has implications for how the simulations are conditioned during the management strategy testing. While it is still the case tha tthe general properties of each SMU are used to set the scene for each SMU's component populations, the advent of sufficient GPS logger data now allows a more formal definition of the productivity of each population. Currently, the modelling assumes that each population is linearly arranged around the coastline, and this will allow, where available, specific catch/yield data to be used to define the bounds of each population within an SMU. 

An important change from previous designs is that larval dispersal is implemented explicitly rather than implicitly being held constant. This alters the dynamics so that analytical equilibrium methods no longer work and we need to resort to iterative approaches to equilibrium if the larval dispersal rate is > 0.0.

Of course, adding such a component increases the number of options that may need to be explored, but currently we envisage including a very low, a middle range, and a relatively high level of larval dispersal, to examin its implications.

## Installation

You can install the development version from [GitHub](https://github.com/haddonm/aMSE) with:

``` r
# install.packages("devtools")
devtools::install_github("haddonm/aMSE")
devtools::install_github(haddonm/rutilsMH) # also useful for plotting, etc
```
## Example

This is a basic example which illustrates the generation of an initial equilibrium. It uses built in data-sets but usually you would generate regionC (constants) and regionD (the dynamic components) from the conditioning data:

```{r example}
# basic example code
library(aMSE)
library(rutilsMH)

  data(region1)
  glb <- region1$globals
  data(testregC)
  data(testregD)
  regDe <- testequil(regC=testregC,regD=testregD,glob=glb)
  str(regDe)
  
```

