---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# LATEST UPDATE

* 2021_07-28 aMSE 0.0.0.800 Added functions (changecolumn, getrecdevcolumn, gettasdevssq) to assist in automated operating model conditioning on the fishery. Also modified the biology_plots and oneyear and oneyearcat to assist with speeding the processes. Lots of additional minor changes but some important ones to do_MSE (read the help, ?do_MSE).

* 2021-07-20 aMSE 0.0.0.900 Added do_condition and compareCPUE. Both used to speed the conditioning of the operating model, though a number of developments are still under development. Currently can automatically search for the AvRec value that will optimize the sum-of-squared residual fit between the
observed CPUE in the historical period, and those predicted by the conditioned model. The undeveloped bit related to the ad hoc recruitment deviates.

# aMSE

<!-- badges: start -->
<!-- badges: end -->

This packages up a new Abalone Management Strategy Evaluation framework. It has a novel structure that is based around the spatial ideas of populations within spatial management units (SAUs), within a zone:

* zone - highest geographical level. Is simply the totality of the spatial management units.

* SAU - spatial assessment units. In Tasmania these would currently be the classical statistical blocks.

* population - literally a population. These are the active components within the simulation framework. The dynamics of the simulation are based around the populations, although, with positive larval dispersal (the default) there is some dependency of neighbouring populations on each other.


This conceptual structure has implications for how the simulations are conditioned during the management strategy testing. While it is still the case that the general properties of each SAU are used to set the scene for each SAU's component populations, the advent of sufficient GPS logger data now allows a more formal definition of the productivity of each population. Currently, the modelling assumes that each population is linearly arranged around the coastline, and this will allow, where available, specific catch/yield data to be used to define the bounds of each population within an SAU. 

An important change from previous designs is that larval dispersal is implemented explicitly rather than implicitly being held constant. This alters the dynamics so that analytical equilibrium methods no longer work and we need to resort to iterative approaches to equilibrium if the larval dispersal rate is > 0.0.

Of course, adding such a component increases the number of options that may need to be explored, but currently we envisage including a very low, a middle range, and a relatively high level of larval dispersal, to examine its implications.

## Installation

Once the MSE framework is made public you can install the development version from [GitHub](https://github.com/haddonm/aMSE) with:

```{r echo=TRUE, eval=FALSE}
if (!require(devtools)){install.packages("devtools")} 

devtools::install_github("https://github.com/haddonm/aMSE",build_vignettes = TRUE)
```

Alternatively, while the development version remains private, you can generate a branch that you can work on by cloning the repository, which, again, can be done very simply within RStudio. Open the New Project option in the project dialog at the top right of the RStudio screen and selection Version Control, then use 'https://github.com/haddonm/aMSE' in the top box, identify where you want the new directory put, and press return. ALternatively, you could download the zip file from inside the 'code' button and establish an R project from that.

It would be a good idea to read Hadley Wickham's draft chapter on Git and GitHub at https://r-pkgs.org/index.html.


## Example

This is an example which illustrates the generation of an initial equilibrium, which then goes on to apply an early version of the Tasmania MCDA using only 50 replicates (that bit takes about 15 seconds on my computer, hence 2.5 minutes for 1000). It uses built in data-sets but usually you would read in a control file, which would contain the name of the biological datafile describing each population.

```{r example, echo=TRUE}
# a constant TAC example
starttime <- (Sys.time())
library(aMSE)
library(rutilsMH)
library(makehtml)
library(knitr)
# OBVIOUSLY you should modify the rundir and datadir to suit your own computer
if (dir.exists("c:/Users/User/DropBox")) {
  ddir <- "c:/Users/User/DropBox/A_codeUse/aMSEUse/scenarios/"
} else {
  ddir <- "c:/Users/Malcolm/DropBox/A_codeUse/aMSEUse/scenarios/"
}
doproject <- TRUE  # change to FALSE if only conditioning is required
verbose <- TRUE
postdir <- "testnew"  # also used to label the output HTML file 
rundir <- paste0(ddir,postdir)
datadir <- paste0(ddir,"tasdata")
alldirExists(rundir,datadir,verbose=verbose)
source(paste0(datadir,"/TasmanianHS.R"))
controlfile <- "controlsau.csv"
# equilibrium zone -------------------------------------------------------------
# You now need to ensure that there is, at least, a control.csv, and a 
# constantsdata.csv file in the data directory plus some other data .csv files
# depending on how conditioned you want the model to be. Templates for the
# correct format can be produced using ctrlfiletemplate(), datafiletemplate().
# 
# Of course, usually one would use data files, control.csv and a zone.csv, which
# is listed as the datafile within the control.csv. These must be stored in 
# rundir. There are example control and data files in the DropBox folder:
# C:\Users\User\Dropbox\National abalone MSE\aMSE_files. Copy the control2.csv
# and zonewest.csv into you rundir. Then, assuming yo have the very latest
# version of aMSE = 5300, the following code should work. Recently both aMSE and
# makehtml have been altered so that all references to resdir have been changed
# to rundir, so both packages will need updating. See their respective GitHub
# readme pages for details.

# HarvestStrategy.R should be in datadir, if it is to be shared by scenarios
source(paste0(datadir,"/TasmanianHS.R"))
# run the scenario --------------------- Obviously unhash this to make it work
# out <- do_MSE(rundir,controlfile,datadir,hsargs=hsargs,hcrfun=mcdahcr,
#                  sampleCE=tasCPUE,sampleFIS=tasFIS,sampleNaS=tasNaS,
#                  getdata=tasdata,calcpopC=calcexpectpopC,varyrs=7,startyr=50,
#                  cleanslate=TRUE,verbose=TRUE,doproject=doproject,ndiagprojs=4,
#                  openfile=TRUE,savesauout=FALSE)
# make results webpage ---------------------------------------------------------
# replist <- list(starttime=as.character(out$starttime),
#                 endtime=as.character(out$projtime))
# glb <- out$glb
# projy <- ifelse(doproject,glb$pyrs,0)
# runnotes <- paste0(out$ctrl$runlabel,":  RunTime = ",out$tottime,
#                    "  replicates = ",glb$reps,",   years projected = ",projy,
#                    "  Populations = ",glb$numpop," and SAU = ",glb$nSAU,
#                    "  Randomseed for conditioning = ",out$ctrl$randseed)
# 
# make_html(replist = replist,  rundir = rundir,  width = 500,  openfile = TRUE,
#           runnotes = runnotes,   verbose = FALSE,  packagename = "aMSE",
#           htmlname = postdir)


```

After running the whole, even if you do not generate the results webpage, you could also try:

```{r }
#str(out$zoneDP,max.level=1)
#str1(out$hcrout)

```

To see the structure of the dynamic object generated by the projections.

See the vignette Running_aMSE.Rmd for a more detailed example.

See the New.Rd for recent developments
