---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# aMSE

<!-- badges: start -->
<!-- badges: end -->

This packages up a new Abalone Management Strategy Evaluation framework. It has a novel structure that is based around the spatial ideas of populations within spatial management units (SAUs), within a zone:

* zone - highest geogrpahical level. Is simply the totality of the spatial management units.

* SAU - spatial assessment units. In Tasmania these would currently be the classical statistical blocks.

* population - literally a population. These are the active components within the simulation framework. The dynamics of the simulation are based around the populations, although, with positive larval dispersal (the default) there is some dependency of neighbouring populations on each other.


This conceptual structure has implications for how the simulations are conditioned during the management strategy testing. While it is still the case that the general properties of each SAU are used to set the scene for each SAU's component populations, the advent of sufficient GPS logger data now allows a more formal definition of the productivity of each population. Currently, the modelling assumes that each population is linearly arranged around the coastline, and this will allow, where available, specific catch/yield data to be used to define the bounds of each population within an SAU. 

An important change from previous designs is that larval dispersal is implemented explicitly rather than implicitly being held constant. This alters the dynamics so that analytical equilibrium methods no longer work and we need to resort to iterative approaches to equilibrium if the larval dispersal rate is > 0.0.

Of course, adding such a component increases the number of options that may need to be explored, but currently we envisage including a very low, a middle range, and a relatively high level of larval dispersal, to examine its implications.

## Installation

Once the MSE framework is made public you can install the development version from [GitHub](https://github.com/haddonm/aMSE) with:

```{r echo=TRUE, eval=FALSE}
if (!require(devtools)){install.packages("devtools")} 

devtools::install_github("https://github.com/haddonm/aMSE",build_vignettes = TRUE)
```

Alternatively, while the development version remains private, you can generate a branch that you can work on by cloning the repository, which, again, can be done very simply within RStudio. Open the New Project option in the project dialog at the top right of the RStudio screen and selection Version Control, then use 'https://github.com/haddonm/aMSE' in the top box, identify where you want the new directory put, and press return. ALternatively, you could download the zip file from inside the 'code' button and establish an R project from that.

It would be a good idea to read Hadley Wickham's draft chapter on Git and GitHub at https://r-pkgs.org/index.html.


## Example

This is an example which illustrates the generation of an initial equilibrium, which then goes on to apply an early version of the Tasmania MCDA using only 100 replicates (that bit takes about 15 seconds on my computer, hence 2.5 minutes for 1000). It uses built in data-sets but usually you would read in a control file, which would contain the name of the biological datafile describing each population.

```{r example, echo=TRUE}
# a constant TAC example
starttime <- (Sys.time())
library(aMSE)
library(rutilsMH)
library(makehtml)
library(knitr)
# Obviously you should modify the resdir to suit your own computer
if (dir.exists("c:/Users/User/DropBox")) {
  ddir <- "c:/Users/User/DropBox/A_code/"
} else {
  ddir <- "c:/Users/Malcolm/DropBox/A_code/"
}
resdir <- paste0(ddir,"aMSEUse/conddata/generic")  # data and results directory
dirExists(resdir,make=TRUE,verbose=TRUE)
# equilibrium zone -------------------------------------------------------------
# You now need to ensure that there is, at least, a control.csv, and a 
# constantsdata.csv file in the data directory plus some other data .csv files
# depending on how conditioned you want the model to be. Templates for the
# correct format can be produced using ctrlfiletemplate(), datafiletemplate().
# In the meantime it is easier to use the included data file "zone".
data(zone)
# Of course, usually one would use data files, control.csv and a zone.csv, which is 
# listed as the datafile within the control.csv. These must be stored in resdir
# zone <- makeequilzone(resdir,"control2.csv") # normally would read in a file
    equiltime <- (Sys.time())   # let's change the initial depletion
    origdepl <-  c(0.30,0.31,0.29,0.32,0.30,0.31,0.29,0.32) 
zoneDD <- depleteSAU(zone$zoneC,zone$zoneD,zone$glb,origdepl,zone$product,len=12)
out <- prepareprojection(zone$zone1,zone$zoneC,zone$glb,zoneDD,zone$ctrl)
  zoneDR <- out$zoneDP
  projC <- out$projC
  zoneCP <- out$zoneC
  midtime <- (Sys.time())

  glb <- zone$glb
  ctrl <- zone$ctrl
  print(equiltime - starttime)
  print(midtime - equiltime)
  propD <- getzoneprops(zone$zoneC,zoneDD,glb,year=1)
  # round(propD,3)
# Do the replicates ------------------------------------------------------------
  inityr <- zone$zone1$projC$inityrs
  saunames <- zone$zone1$SAUnames
  sauindex <- zone$glb$sauindex
  pyrs <- projC$projyrs + projC$inityr
  B0 <- tapply(sapply(zone$zoneC,"[[","B0"),sauindex,sum)
  exB0 <- tapply(sapply(zone$zoneC,"[[","ExB0"),sauindex,sum)

  midtime <- (Sys.time())
mseproj <- doprojection(zoneCP,zoneDR,glb,ctrl,projC$projyrs,applyHS=mcdahcr,projC$inityrs)
sauzoneDP <- asSAU(mseproj,sauindex,saunames,B0,exB0)
  endtime <- (Sys.time())
  print(endtime - midtime)
# Now plot some results
  plotC <- function(nsau,saunames,reps,projyrs,plts=c(4,2)) {
    return(list(nsau=nsau,saunames=saunames,reps=reps,projyrs=projyrs,plts=plts))
  }

  saunames <- zone$zone1$SAUnames
  pyrs <- projC$projyrs + projC$inityr
  reps <- ctrl$reps
  label <- "Catches t"
  yrs <- 1:pyrs

plotconst <- plotC(nsau=length(saunames),saunames,reps,pyrs,plts=c(4,2))

plotprep(width=7,height=8,newdev=FALSE)
plotproj(sauzoneDP$cpue,"CPUE",plotconst,vline=projC$inityrs,addqnts=TRUE)

plotproj(sauzoneDP$saudeplsB,"Spawning Biomass Depletion",plotconst,vline=projC$inityrs,addqnts=TRUE)

plotproj(sauzoneDP$recS/1000,"Recruitment '000s",plotconst,vline=projC$inityrs,addqnts=TRUE)

plotproj(sauzoneDP$catS,"Catches t",plotconst,vline=projC$inityrs,addqnts=TRUE)

```


See the vignette Running_aMSE.Rmd for a more detailed example.

See the New.Rd for recent developments
