---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# LATEST UPDATE

* 2021-04-14 0.0.0.3200 Added NumNe to zoneDP output ready for its use in estimating FIS results by population and other numbers-at-size related indices. NumNe are the numbers-at-size following growth and the application of half of natural mortality but before any fishing mortality has occurred. (see News.ms for history of development) 

# aMSE

<!-- badges: start -->
<!-- badges: end -->

This packages up a new Abalone Management Strategy Evaluation framework. It has a novel structure that is based around the spatial ideas of populations within spatial management units (SAUs), within a zone:

* zone - highest geogrpahical level. Is simply the totality of the spatial management units.

* SAU - spatial assessment units. In Tasmania these would currently be the classical statistical blocks.

* population - literally a population. These are the active components within the simulation framework. The dynamics of the simulation are based around the populations, although, with positive larval dispersal (the default) there is some dependency of neighbouring populations on each other.


This conceptual structure has implications for how the simulations are conditioned during the management strategy testing. While it is still the case that the general properties of each SAU are used to set the scene for each SAU's component populations, the advent of sufficient GPS logger data now allows a more formal definition of the productivity of each population. Currently, the modelling assumes that each population is linearly arranged around the coastline, and this will allow, where available, specific catch/yield data to be used to define the bounds of each population within an SAU. 

An important change from previous designs is that larval dispersal is implemented explicitly rather than implicitly being held constant. This alters the dynamics so that analytical equilibrium methods no longer work and we need to resort to iterative approaches to equilibrium if the larval dispersal rate is > 0.0.

Of course, adding such a component increases the number of options that may need to be explored, but currently we envisage including a very low, a middle range, and a relatively high level of larval dispersal, to examine its implications.

## Installation

Once the MSE framework is made public you can install the development version from [GitHub](https://github.com/haddonm/aMSE) with:

```{r echo=TRUE, eval=FALSE}
if (!require(devtools)){install.packages("devtools")} 

devtools::install_github("https://github.com/haddonm/aMSE",build_vignettes = TRUE)
```

Alternatively, while the development version remains private, you can generate a branch that you can work on by cloning the repository, which, again, can be done very simply within RStudio. Open the New Project option in the project dialog at the top right of the RStudio screen and selection Version Control, then use 'https://github.com/haddonm/aMSE' in the top box, identify where you want the new directory put, and press return. ALternatively, you could download the zip file from inside the 'code' button and establish an R project from that.

It would be a good idea to read Hadley Wickham's draft chapter on Git and GitHub at https://r-pkgs.org/index.html.


## Example

This is an example which illustrates the generation of an initial equilibrium, which then goes on to apply an early version of the Tasmania MCDA using only 100 replicates (that bit takes about 15 seconds on my computer, hence 2.5 minutes for 1000). It uses built in data-sets but usually you would read in a control file, which would contain the name of the biological datafile describing each population.

```{r example, echo=TRUE}
# a constant TAC example
starttime <- (Sys.time())
library(aMSE)
library(rutilsMH)
library(makehtml)
library(knitr)
# OBVIOUSLY you should modify the rundir and datadir to suit your own computer
if (dir.exists("c:/Users/User/DropBox")) {
  ddir <- "c:/Users/User/DropBox/A_code/"
} else {
  ddir <- "c:/Users/Malcolm/DropBox/A_code/"
}
rundir <- paste0(ddir,"aMSEUse/scenarios/HS652510")
datadir <- paste0(ddir,"aMSEUse/scenarios/tasdata")
alldirExists(rundir,datadir)
# equilibrium zone -------------------------------------------------------------
# You now need to ensure that there is, at least, a control.csv, and a 
# constantsdata.csv file in the data directory plus some other data .csv files
# depending on how conditioned you want the model to be. Templates for the
# correct format can be produced using ctrlfiletemplate(), datafiletemplate().
# 
# Of course, usually one would use data files, control.csv and a zone.csv, which
# is listed as the datafile within the control.csv. These must be stored in 
# rundir. There are example control and data files in the DropBox folder:
# C:\Users\User\Dropbox\National abalone MSE\aMSE_files. Copy the control2.csv
# and zonewest.csv into you rundir. Then, assuming yo have the very latest
# version of aMSE = 5300, the following code should work. Recently both aMSE and
# makehtml have been altered so that all references to resdir have been changed
# to rundir, so both packages will need updating. See their respective GitHub
# readme pages for details.

# TasmanianHS.R should be in rundir, but during development is in data-raw
source(paste0(rundir,"/TasmanianHS.R"))
# generate equilibrium zone ----------------------------------------------------
starttime2 <- (Sys.time())
zone <- makeequilzone(rundir,"controlsau.csv",datadir=datadir,cleanslate = TRUE)
equiltime <- (Sys.time()); print(equiltime - starttime2)
# declare main objects ---------------------------------------------------------
glb <- zone$glb
ctrl <- zone$ctrl
zone1 <- zone$zone1
projC <- zone1$projC
condC <- zone1$condC
zoneC <- zone$zoneC
zoneD <- zone$zoneD
product <- zone$product
# add some tables and plots into rundir
biology_plots(rundir, glb, zoneC)
plotproductivity(rundir,product,glb)
numbersatsize(rundir, glb, zoneD)
# Condition on Fishery ---------------------------------------------------------
zoneDD <- dohistoricC(zoneD,zoneC,glob=glb,condC,sigR=1e-08,sigB=1e-08)

propD <- t(getzoneprops(zoneC,zoneDD,glb,year=47))
addtable(round(propD,4),"propertyDD.csv",rundir,category="zoneDD",caption=
           "Properties of zoneD after conditioning on historical catches.")
addtable(round(t(zoneDD$harvestR[40:47,]),4),"final_harvestR.csv",rundir,
         category="zoneDD",caption="Last eight years of harvest rate.")
popdefs <- getlistvar(zone$zoneC,"popdef")
addtable(round(t(popdefs),3),"popdefs.csv",rundir,category="zoneDD",caption=
           "Population specific definitions")
# Prepare projections ----------------------------------------------------------
plotconditioning(zoneDD,glb,zoneC,condC$histCE,1973:2019,rundir)
cmcda <- mcdahcr(arrce=condC$histCE,hsargs=hsargs,
                 yearnames=rownames(condC$histCE),saunames=glb$saunames,
                 acatches=tail(condC$histCatch,1))
pms <- cmcda$pms
multTAC <- cmcda$multTAC
out <- prepareprojectionnew(projC=projC,condC=condC,zoneC=zoneC,glb=glb,
                            multTAC=multTAC,zoneDD=zoneDD,ctrl=ctrl,varyrs=7,
                            lastsigR = ctrl$withsigR)
zoneDP <- out$zoneDP
projC <- out$projC
zoneCP <- out$zoneCP
zoneDDR <- out$zoneDDR

zoneDP <- doTASprojections(ctrl,zoneDP,zoneCP,condC$histCE,glb,mcdahcr,hsargs)
# save more plots to rundir ----------------------------------------------------
histCE <- condC$histCE
B0 <- getvar(zoneC,"B0")
ExB0 <- getvar(zoneC,"ExB0")
sauCI <- sauplots(zoneDP,zoneDDR,glb,rundir,B0,ExB0,startyr=25,addCI=TRUE,
                  histCE=histCE)
outzone <- poptozone(zoneDP,glb,
                     B0=sum(getvar(zoneC,"B0")),
                     ExB0=sum(getvar(zoneC,"ExB0")))
plotZone(outzone,rundir,CIprobs=c(0.05,0.5,0.95),addfile=TRUE)
projtime <- Sys.time()
print(projtime - starttime)
# make results webpage ---------------------------------------------------------
replist <- list(starttime=as.character(starttime),endtime=as.character(projtime))

# Unhash the make_html command to generate the results webspage
# make_html(
#   replist = replist,
#   rundir = rundir,
#   width = 500,
#   openfile = TRUE,
#   runnotes = NULL,
#   verbose = FALSE,
#   packagename = "aMSE",
#   htmlname = "aMSE"
# )


```

After running the whole, even if you do not generate the results webpage, you could also try:

```{r }
str(zoneDP,max.level=1)

```

To see the structure of the dynamic object generated by the projections.

See the vignette Running_aMSE.Rmd for a more detailed example.

See the New.Rd for recent developments
